Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-06-12 GWLDCI.tex
Paper 1 edited up to section 4

In recent years, there has been an active program researching the geometry and combinatorics underlying SYM $N=4$ theory \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, AmplituhedronDecomposition, galashinlam}. This body of work started with the observation that the on shell amplitudes\note{Is this for tree level amplitudes?  If so, we should say that} of this theory correspond to the volume of a subspace of a positive Grasmannian, called an Amplituhedron \cite{Arkani-Hamed:2013jha}. Since then, there has been significant work studying the structure of the amplituhedron, geometrically and combinatorially, for instance \cite{Arkani-Hamed:2013kca, arkani:2012nw, AmplituhedronDecomposition, galashinlam}. The focus of much of this work has been on understanding the relationship between the combinatorial objects from the physics of the system, BCFW diagrams, and their corresponding natural object for studying the positive Grassmannians, positroid cells.

@@

The main result of Section~\ref{sec equiv} is Theorem~\ref{same matroid iff equiv}, which says that two Wilson loop diagrams define the same matroid if and only if they are equivalent.  In subsection~\ref{sec: polygon partitions} we relate certain subdiagrams (the {\em exact} subdiagrams, Definition~\ref{def:exact diagram}) of Wilson loop diagrams to triangulations of polygons, and then in subsection~\ref{sec: exact diagram matroidal props} we examine the matroidal properties of the exact subdiagrams.  Specifically, we show that a subdiagram of a Wilson loop diagrams $W$ defines a uniform matroid if and only if the subdiagram is exact (Theorem \ref{exactuniformthm}), and that this uniform matroid \hlfix{is isomorphic to}{isomorphic or equal?} the dual matriod of $W$ restricted to the exact subdiagram (Remark \ref{remark exact dual restiction}).  Subsection~\ref{sec: matroids and equivalence} is devoted to the proof of Theorem~\ref{same matroid iff equiv}, along with an easy corollary which gives a formula for the size of each equivalence class (Corollary~\ref{number of equiv diagrams}).

@@

\begin{rmk}\label{alt F(P) rmk}
An equivalent definition of $F(P)$ is \bas F(P)  = \big(V(P) \setminus V(P^c) \big) \cup \{\text{ all non-supporting vertices in $W$ }\} \;,\eas 
\note{replaced $F(\emptyset)$ with words in this definition because we were defining $F(P)$ in terms of $F(\emptyset)$ and it was mildly inelegant/recursive)}
i.e. $F(P)$ can be thought of as the set of vertices that either only support propagators in $P$ or do not support any propagators at all. Furthermore, note that by construction $\Prop(F(P)) \subseteq P$.  This containment is strict if there is any propagator $p$ in $P$ with the property that every vertex in the support of $p$ also supports a propagator not in $P$.
\end{rmk}

@@

In other words, a Wilson loop diagram, $(\cP, [n])$ is admissible if and only if $n \geq |\cP| +4$, and has no crossing propagators, no propagators of the form $p = (i, i+1)$, and no pairs of propagators that start and end on the same pair of non-adjacent edges.
\note{My concern here is with this last sentence.  I think it is in fact true, but I don't see that it is obvious from the definitions.  It is saying that if any set of propagators is too dense according to condition 2, then either a single propagator in that set or a pair of them is already too dense (or condition 3 fails).  That seems nontrivial to me, even interesting.  Probably you proved this in your first paper on the topic and it is so old hat you don't even think about it any more, but in that case the appropriate lemma should be cited, rather than implying it is just a rephrasing of the definition itself}

@@

These $C(W)$ parametrize a subspace of $\Gr(k, n)$ as shown in \cite{wilsonloop}, call it $\Sigma(W)$.
\note{I don't think this sentence stands properly alone: What set precisely is ``These'' running over?  $\Sigma(W)$ is not the whole subspace of $\Gr(k,n)$ it is just the part corresponding to $C(W)$, so the obvious ``it'' in the sentence (namely the subspace) is not the correct one.  Also, $\Sigma(W)$ appears only one other time in this paper (in a more formal definiiton later which is essentially saying the same thing as this sentence but in more detail), perhaps we should refer forwards; perhaps we don't need this notation in this paper?}

@@

In particular, this allows us to prove results about $M(W)$ by considering the behavior of propagators in $W$. We record a few elementary facts about the rank and cycles of $M(W)$ here as an example of this.

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle in $M(W)$, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support disjoint sets of propagators and in such a way that we obtain a partition of the propagator set, i.e. if we have \bas [n] = \sqcup_i V(P_i) \quad \textrm{ such that } \quad \sqcup P_i = \cP;  \text{ with }V(P_i) \cap V(P_j) = \emptyset \text{ and } P_i \cap P_j = \emptyset \quad \forall i \neq j\;, \eas then the matroid $M(W)$ is separable, i.e. \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\item $F(P)$ is a flat of $M( W)$, which we call the \emph{propagator flat} of $P$.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}. (3) is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ (from the definition of $F(P)$) and no subset of $S$ supports fewer propagators than the number of vertices it contains (by Theorem~\ref{thm WLD defines matroid} this is one characterization of an independent set in $M(W)$). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup\{v\}$ also satisfies this independence condition. Thus $\rk(F(P) \cup\{v\}) = \rk(F(P)) + 1$, as required.
\end{proof}

@@

\begin{rmk} \label{remark exact dual restiction} For readers familiar with duals and complements of matroids, we reinterpret Theorem~\ref{exact diagrams contractions} in these terms. If $M^*$ denotes the dual of a matroid, then by standard matroid theory \hlfix{[ref]}{if we're not going to define them here (and we really shouldn't!) then it's probably polite to give a reference} the dual of a contraction of a matroid is the same as the restriction of the dual matroid by the complement: \bas M / S = M^*|_{S^c} \;.\eas Then Theorem \ref{exact diagrams contractions} implies that, for $(P, V(P))$ an exact subdiagram of $W$, we have \bas M\big((P, V(P))\big) = M(W)  / F(P^c) = M(W)^*|_{V(P)} \;.\eas
\end{rmk}

@@

\begin{dfn}
Given a Wilson loop diagram $W = (\cP, [n])$ and its associated positroid $M(W)$, let $\Sigma(W)$ denote the corresponding positroid cell in the CW complex on $\Gr(|\cP|, n)$.
\end{dfn}

@@

\begin{proof}
The unique top dimensional cell of $\Gr(|P|, |V(P)|)$ is defined by precisely those points in $\Gr(|P|, |V(P)|)$ with the property that all Pl\"ucker coordinates are strictly greater than $0$. Since $(P, V(P))$ is an exact subdiagram, the matroid is uniform, and so all $|P| \times |P|$ minors are non-zero. Intersecting these with the cases with the positive Grassmannians demands that all minors be strictly positive.\note{I don't understand this last sentence.}
\end{proof}

@@

Assume that $W$ and $W'$ are equivalent. Without loss of generality, write $W = (P \cup R, [n])$ and $W' = (P \cup R', [n])$, where $P \subset \cP \cap \cP'$ and $(R, V(R))$ and $(R', V(R'))$ are two maximal exact subdiagrams, with $R \neq R'$, but $V(R) = V(R')$. If this is not the case, one may always find a family of diagram, $\{W_i\}$ satisfying this condition and forming a transitive chain connecting $W$ to $W'$ in the equivalence class.

Let $U \subset V(R)$ be any subset of size $|U| = |R|$. Since $(R, V(R))$ defines a uniform matroid, by Lemma \ref{exactuniformthm}, the set $U$ is independent in the subdiagram $(R, V(R))$, and thus in $W$. The complementary set $F(R^c) = F(P)$ is a flat of maximal rank by lemma \ref{maxexactcomplementrank} ($\rk(F(P)) = |P|$). Let $B \subset F(P)$ be a maximal independent set ($\rk B = |P|$) in $F(P)$. Since $F(P)$ is a flat, adding any element of $V(R)$ to $B$ increases the rank. Therefore, any basis of $W$ can be written as $B \cup U$ for some $B$ and $U$ of the form indicated. However, since $F(P)$ is common to both $W$ and $W'$, and $V(R) = V(R')$, any any basis of $W'$ can also be written $B \cup U$. Thus both matroids have the same bases sets, proving that they are the same.

For the converse, assume that the matroids associated to $W$ and $W'$ are the same: $M(W) = M(W')= M$. Let $\mathcal{R} = \{(P_i, V(P_i)\}_{i=1}^k$ and $\mathcal{R}' = \{(P'_i, V(P'_i)\}_{i=1}^l$ be the sets of maximal exact subdiagrams of $W$ and $W'$. Write $F_i = F(P_i^c)$ and $F'_i = F(P^{'c}_i)$ to be the complementary cyclic flats. By Theorem \ref{exactuniformthm} and Theorem \ref{exact diagrams contractions},  $M/F_i$ and $M/F_i'$ are uniform matroids. Since, by Theorem \ref{exactuniformthm}, any subset $V$ of $M$ such that $M/(V^c)$ is uniform defines an exact subdiagram of both $W$ and $W'$. If $U \subset V$ are two such sets, then the corresponding exact subdiagrams are also subsets. Therefore, $|\mathcal{R}| = |\mathcal{R}'|$, and we write $V(P_i) = V(P'_i)$, after possibly reindexing $\mathcal{R}$ and $\mathcal{R}'$.

Since the sets of propagators defining maximal exact subdiagrams partition $\cP$, by Lemma \ref{uniqueproppartitioncor}, write $\cup_{i = 1}^k P_i = \cup_{i = 1}^k P'_i = \cP$. Reorganize the vertex sets of maximal exact subdiagrams as follows: \ba \cup_{P_i \not \in \mathcal{R}'} V(P_i) = V(\cup_{P_i \not \in \mathcal{R}'} P_i)  = F(\cup_{P_i \in \mathcal{R}'} P_i)^c\label{vertexsets}\; .\ea The final flat may, of course, be empty.

Thus, we have partitioned the vertices of $W$ and $W'$ into two complementary sets. The first, $ \cup_{P_i \not \in \mathcal{R}'} V(P_i)$,  is comprised of the union of supports of maximal exact subdiagrams whose propagators differ between $W$ and $W'$. The second, $ F(\cup_{P_i \in \mathcal{R}'} P_i)$,  is the propagator flat of all propagators in common between $W$ and $W'$.

Without loss of generality, assume that $|\mathcal{R}| = |\mathcal{R}'| = 1$. Then the two Wilson loop diagrams are equivalent. If  $|\mathcal{R}| = |\mathcal{R}'| > 1$, then one may define a family of Wilson loop diagrams, $W_0$ to $W_k$ defined such that $W_0 = W$ and $W_i$ is derived from $W_{i-1}$ by replacing the propagator set $P_i$ with $P'_i$. In this manner, $W' = W_k$ and $W_i \sim W_{i+1}$, making $W \sim W'$.
\end{proof}

\note{Two comments on this proof.  First I think it was written before other things and so doesn't quite match in style and language; I have improved this a bit but it would benefit from more editing with that in mind.  Second, I feel that both directions essentially come down to the fact that there is a unique uniform matroid of each dimension and rank.}

@@

\sanote{I undid Sian's edit here. I like "saying what i'm about to say. Saying it. Summarizing what I said"}\note{Sian: sure. I'll push back if it gets too repetitive, but it works here.}In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.  We do not obtain a formula for this, but do obtain an alternate characterization of the equvalence classes which is the subject of the next section.

>>>>

In recent years, there has been an active program researching the geometry and combinatorics underlying SYM $N=4$ theory \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, AmplituhedronDecomposition, galashinlam}. This body of work started with the observation that the on shell (tree level) amplitudes of this theory correspond to the volume of a subspace of a positive Grasmannian, called an Amplituhedron \cite{Arkani-Hamed:2013jha}. Since then, there has been significant work studying the structure of the amplituhedron, geometrically and combinatorially, for instance \cite{Arkani-Hamed:2013kca, arkani:2012nw, AmplituhedronDecomposition, galashinlam}. The focus of much of this work has been on understanding the relationship between the combinatorial objects from the physics of the system, BCFW diagrams, and their corresponding natural object for studying the positive Grassmannians, positroid cells.

@@

The main result of Section~\ref{sec equiv} is Theorem~\ref{same matroid iff equiv}, which says that two Wilson loop diagrams define the same matroid if and only if they are equivalent.  In subsection~\ref{sec: polygon partitions} we relate certain subdiagrams (the {\em exact} subdiagrams, Definition~\ref{def:exact diagram}) of Wilson loop diagrams to triangulations of polygons, and then in subsection~\ref{sec: exact diagram matroidal props} we examine the matroidal properties of the exact subdiagrams.  Specifically, we show that a subdiagram of a Wilson loop diagrams $W$ defines a uniform matroid if and only if the subdiagram is exact (Theorem \ref{exactuniformthm}), and that this uniform matroid is equal to the dual of the matriod associated to $W$ restricted to the exact subdiagram (Remark \ref{remark exact dual restiction}).  Subsection~\ref{sec: matroids and equivalence} is devoted to the proof of Theorem~\ref{same matroid iff equiv}, along with an easy corollary which gives a formula for the size of each equivalence class (Corollary~\ref{number of equiv diagrams}).

@@

\begin{rmk}\label{alt F(P) rmk}
An equivalent definition of $F(P)$ is \bas F(P)  = \big(V(P) \setminus V(P^c) \big) \cup \{\text{ all non-supporting vertices in $W$ }\} \;,\eas 
i.e. $F(P)$ can be thought of as the set of vertices that either only support propagators in $P$ or do not support any propagators at all. Furthermore, note that by construction $\Prop(F(P)) \subseteq P$.  This containment is strict if there is any propagator $p$ in $P$ with the property that every vertex in the support of $p$ also supports a propagator not in $P$.
\end{rmk}

@@

In other words, a Wilson loop diagram, $(\cP, [n])$ is admissible only if $n \geq |\cP| +4$, and has no crossing propagators, no propagators of the form $p = (i, i+1)$, and no pairs of propagators that start and end on the same pair of non-adjacent edges.

{\color{violet} INTERNAL NOTE ONLY. (I've removed the offending "if" above: The other direction of the above is best seen from the point of view of triangulations. Any exact subdiagram corresponds to a triangulation. $n>1$ or more propagators forms a minimal violation the the density condition if one can remove one propagator to get an exact subdiagram on $n-1$ propagators, and thus a triangulation. The only way this can happen is to cross propagators/ diagonals, or two have the diagonal start and end at the same vertex (parallell propagators). A single violator violates the density condition if it starts and ends on adjascent edges of the Wilson loop diagram/ vertices of the triangulation.}

@@

@@

In particular, this allows us to prove results about $M(W)$ by considering the behavior of propagators in $W$. We record a few facts about the ranks cycles and flats of $M(W)$ here as an example of this.

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle in $M(W)$, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support disjoint sets of propagators and in such a way that we obtain a partition of the propagator set, i.e. if we have \bas [n] = \sqcup_i V(P_i) \quad \textrm{ such that } \quad \sqcup P_i = \cP;  \text{ with }V(P_i) \cap V(P_j) = \emptyset \text{ and } P_i \cap P_j = \emptyset \quad \forall i \neq j\;, \eas then the matroid $M(W)$ is separable, i.e. \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\item $F(P)$ is a flat of $M( W)$, which we call the \emph{propagator flat} of $P$.
\item If $F$ is a cyclic flat of $M(W)$, then it is a propagator flat.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}. (3) is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ (from the definition of $F(P)$) and no subset of $S$ supports fewer propagators than the number of vertices it contains (by Theorem~\ref{thm WLD defines matroid} this is one characterization of an independent set in $M(W)$). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup\{v\}$ also satisfies this independence condition. Thus $\rk(F(P) \cup\{v\}) = \rk(F(P)) + 1$, as required.

Point (5) is given in \cite[Lemma 3.28]{wilsonloop}.
\end{proof}

@@

\begin{rmk} \label{remark exact dual restiction} For readers familiar with duals and complements of matroids, we reinterpret Theorem~\ref{exact diagrams contractions} in these terms. If $M^*$ denotes the dual of a matroid, then by standard matroid theory \cite{OxleyMatroidBook} the dual of a contraction of a matroid is the same as the restriction of the dual matroid by the complement: \bas M / S = M^*|_{S^c} \;.\eas Then Theorem \ref{exact diagrams contractions} implies that, for $(P, V(P))$ an exact subdiagram of $W$, we have \bas M\big((P, V(P))\big) = M(W)  / F(P^c) = M(W)^*|_{V(P)} \;.\eas
\end{rmk}

@@

\begin{dfn}
Given a Wilson loop diagram $W = (\cP, [n])$, its associated matroid $M(W)$ corresponds to positroid cells in the CW complex on $\Gr(|\cP|, n)$.
\end{dfn}

@@

\begin{proof}
The unique top dimensional cell of $\Gr(|P|, |V(P)|)$ is defined by precisely those points in $\Gr(|P|, |V(P)|)$ with the property that all Pl\"ucker coordinates are strictly greater than $0$. Since $(P, V(P))$ is an exact subdiagram, the matroid is uniform, and so all $|P| \times |P|$ minors are non-zero. Since $(P, V(P))$ is weakly admissible, the matroid associated to it can be realized by an element of the totally nonnegative Grassmannian $\Gr(k,n)$, by Theorem \ref{thm WLD defines matroid}.
\end{proof}

@@

Assume that $W$ and $W'$ are equivalent. Without loss of generality, write $W = (P \sqcup R, [n])$ and $W' = (P \sqcup R', [n])$, where $P \subset \cP \cap \cP'$ and $(R, V(R))$ and $(R', V(R'))$ are two maximal exact subdiagrams, with $R \neq R'$, but $V(R) = V(R')$. If one cannot find this partition of $\cP$ and $\cP'$, one may always find a family of diagrams, $\{W_1 \ldots W_n\}$ such that $W_{i-1}$ and $W_i$ satisfy this contion, and with $W = W_0$ and $W' = W_{n+1}$. That is, we can find a  transitive chain ($\{W_1 \ldots W_n\}$) between $W$ and $W'$ in the equivalence class.

By Theorem \ref{exactuniformthm}, $(R, V(R))$ defines a uniform matroid. Therefore, any set $U$ of size $|R|$ is independent in the subdiagram $(R, V(R))$, and thus in $W$.

By the third point of Lemma \ref{maxexactcomplementrank}, the complementary flat, $F(P)$ has full rank: $\rk(F(P)) = |P|$. Let $B \subset F(P)$ be a maximal independent subset of $F(P)$: $\rk (B) = |P|$. Since $F(P)$ is a flat, adding any element of $V(R)$ to $B$ increases the rank. Therefore, any basis of $W$ can be written as $B \cup U$ for some $B$ and $U$ of the form indicated. However, since $F(P)$ is common to both $W$ and $W'$, and $V(R) = V(R')$, any any basis of $W'$ can also be written $B \cup U$. Thus both matroids have the same bases sets, proving that they are the same.

For the converse, assume that the matroids associated to $W$ and $W'$ are the same: $M(W) = M(W')= M$. Let $\mathcal{R} = \{(P_i, V(P_i)\}_{i=1}^m$ and $\mathcal{R}' = \{(P'_i, V(P'_i)\}_{i=1}^l$ be the sets of maximal exact subdiagrams of $W$ and $W'$. Recall that these sets $\{P_i\}$ and $\{P_i'\}$ partition the propagator sets $\cP$ and $\cP'$ of $W$ and $W'$ respectivley. Let $\mathcal{F}$ and $\mathcal{F}'$ the respective complementary vertices: \bas \mathcal{F} = \{F_i = V(P_i)^c | (P_i, V(P_i)) \in \mathcal{R}\} \quad  \mathcal{F}' = \{F_i' = V(P'_i)^c | (P'_i, V(P'_i)) \in \mathcal{R}\} \;.\eas By Theorem \ref{maxexactcomplementrank}, every $F_i$ and $F_i'$ is a cyclic flat of full rank in $M$. We claim that $\mathcal{F} = \mathcal{F'}$.

If not, there would be a cyclic flat $F_j$ of $M$. Without loss of generality, write $F_j \in \mathcal{F} \setminus \mathcal{F'}$.  We may write $F_j = V(P_j)^c$, with $(P_j, V(P_j))$ a maximal exact subdiagram in $W$. By Lemma \ref{lem facts about WLD matroids} (5), we may write $F_j= F(Q)$ as a propagator flat. By Lemma \ref{contractsubdiaglem}, $M/F(Q) = F/F_j = M(Q^c, V(Q^c))$, which is uniform, by construction of $F_j$. Theorem \ref{exactuniformthm} $(Q^c, V(Q^c))$ is exact. Since this exact subdiagram is not in $\mathcal{R}'$ by construction of $F_j$, it is not a maximal exact subdiagram of $W'$. That is, there is a maximal exact subdiagram containing it: $(Q^c, V(Q^c)) \subset (P_j', V(P_j'))$. Suppose $|P_j' \setminus Q^c| = m$. By counting vertices of the two nested sets $F_j' \subset F_j$ we see that  \bas |F_j| = n -|Q^c| -3  \textrm{ while } |F_j'| = n -|P_j'| -3 = n -|Q^c| -m -3\;.\eas That is, $F_j'$ has $m$ more vertices than $F_j$. Furthermore, by construction $\rk(F_j) + m = \rk (F_j')$. This violates the assumption that both $F_j$ and $F_j'$ are cyclic flats.

Therefore, we see that $\mathcal{F} = \mathcal{F'}$, which implies that the maximally exact subdiagrams of $W$ and $W'$, are supported on the same vertex sets. All propagators of $W$ and $W'$ are accounted for in these sets. There there exists a transitive chain of equivalent diagrams that put $W$ and $W'$ in the same equivalence class.
\end{proof}

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.  We do not obtain a formula for this, but do obtain an alternate characterization of the equvalence classes which is the subject of the next section.
